---
title: "assignment4"
output: html_document
date: "2022-09-27"
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

The data for this question are stored in the file hw4q1 and correspond
to an experiment to measure the effect of a new drug in the memory of
patients in a nursing home. The patients were tested for memory before
the treatment started and again after one month taking the drug. The
dataset has two variables, mem, the score in the test, and type with two
values, before for the initial score and after for the final score (a)
Load the dataset and check whether type is stored as a factor. If it is
not, transform it into a factor.

```{r}
dataset = read.table("hw4q1") 
str(dataset)
dataset$type = as.factor(dataset$type)
# data  = factor( , levels = "before, after)
str(dataset)
```

(b) Draw boxplots for mem according to type and comment.

```{r}
boxplot(dataset$mem ~dataset$type) # 可以不加前缀 , 可以加颜色 data = data2,col = 'lightblue'
```

after one month taking the drug, the patients obviously have better
score in test. 每个数据集都是symmetric and have similar dispersion.

(c) Draw a scatterplot of the memory score after versus the memory score
    before and comment on what you observe. Do you think the two scores
    are independent?

```{r}
scatter.smooth(dataset$mem[dataset$type=="before"],dataset$mem[dataset$type=="after"])
```

I think the two scores are dependent. Because the polt nearly a straight
line. 答案: 说好听点, 是 linear relation between the scores before and
after treatment.This indicates that these variables are correlated.

(d) We want to determine whether the treatment had an impact on the
    memory score of the patients. State clearly the statistical
    hypothesis that you want to test. What test or tests would you
    consider adequate in this situation and why? What are the
    assumptions? Are they satisfied in this case? Carry out all
    appropriate tests for this problem and comment on your results.

```{r}
t.test(dataset$mem[dataset$type=="before"],dataset$mem[dataset$type=="after"], var.equal = TRUE) 
这是错的!
#  答案如下, 必须要用#paired =True
t.test(data2$mem[data2$type == 'after'],
       data2$mem[data2$type == 'before'],
       paired = TRUE)
#data2$dif <- data2$mem[data2$type == 'after'] -
#  data2$mem[data2$type == 'before']
#qqnorm(data2$dif)
#qqline(data2$dif) # 这个可以检验是正态分布. 因为t
#shapiro.test(data2$dif)
wilcox.test(data2$mem[data2$type == 'after'],
       data2$mem[data2$type == 'before'],
       paired = TRUE)
```

H0: mean1 == mean2 H1: mean1 != mean2 We can use t test. Because n \<30,
std is unknown. We can reject H0, so the treatment actually had an
impact on the memory score of the patients. p很小, 所以拒绝假设.

## Question 2

We will use the data set Pima.te in the MASS package for this question.
Open the help file for this data set and get acquainted with it. We are
going to focus on two variables, bp and type. (a) Divide the plotting
window into two regions, one single column with two rows, and plot
histograms for bp for types Yes and No. Since you want to use these
graphs for comparing the two populations, use the same scales in both
cases. Use reasonable labels for the axes and a title indicating the
corresponding type. Make sure that the area for the figure is large
enough so that the histograms are clearly seen. Compare the two graphs
and comment on similarities and differences.

```{r}
library(MASS)
pim = MASS::Pima.te
str(pim)
par(mfcol=c(2,1))
hist(pim$bp[pim$type=="Yes"],col = 'azure2',xlab='diastolic blood pressure (mm Hg)',xlim=c(20,120),  ylab='Freq',main = "diabetic  ", prob= TRUE) 
hist(pim$bp[pim$type=="No"],col = 'azure2',xlab='diastolic blood pressure (mm Hg)',xlim=c(20,120),  ylab='Freq',main = "No diabetic  ", prob= TRUE)
#答案如下: 
bp.yes <- Pima.te$bp[Pima.te$type == 'Yes']
bp.no <- Pima.te$bp[Pima.te$type == 'No']
par(mfrow=c(2,1))
hist(bp.yes, xlim = c(20,120), ylim = c(0, 0.035), freq = F,
     xlab = 'blood pressure', main = 'Diabetic Pima Women')
hist(bp.no, xlim = c(20,120), ylim = c(0, 0.035), freq = F,
     xlab = 'blood pressure', main = 'Non-Diabetic Pima Women')
```

Similarties: blood pressure mainly in 60-80. Differences: Diabetic
patients have more people over 80 blood pressure.
不过两个dispersion看起来差不多

(b) Boxplot blood pressure as a function of type and comment on the
    graph. Make sure you have a single plotting window with both
    boxplot.

```{r}
boxplot(pim$bp~pim$type)
```

Diabetic people have higher mean blood pressure. boxes高度差不多.

(c) Calculate mean and standard deviation for both types and find how
    many subjects of each type are there in the dataset.

```{r}
mean(pim$bp[pim$type=="Yes"])
sd(pim$bp[pim$type=="Yes"])
#count(pim$bp[pim$type=="Yes"])
mean(pim$bp[pim$type=="No"])
sd(pim$bp[pim$type=="No"])
#count(pim$bp[pim$type=="No"])
table(pim$type)
```

(d) We want to determine if the pulse rate for diabetic women is
    significantly different from a reference value of 70 mm Hg. What
    (parametric) statistical test do you think is appropriate in this
    case? Carry this test out and discuss your results. Describe the
    assumptions you need for this test to be valid and check whether
    they are satisfied by the data set.

```{r}
(tn <- (mean(pim$bp[pim$type=="Yes"])-70)/(sd(pim$bp[pim$type=="Yes"])/sqrt(108)))
t.test(pim$bp[pim$type=="Yes"], mu=70)
qqnorm(bp.yes); qqline(bp.yes)# 证明是正态分布. shapiro.test(bp.yes) 
```

H0: meanbp==70 H1: meanbp!=70

The p value is just below 0.05 and the decision depends on our choice
for α. If we choose 0.05,0.02 or 0.01,the null hypothesis is rejected.
Thus, the pulse rate for diabetic women is significantly different from
a reference value of 70 mm Hg.

拒绝原假设,

(e) We now want to compare the two populations (Pima women with and
    without diabetes) to see if there is a difference in the average
    blood pressure. What (parametric) test would you perform in this
    case? What assumptions are needed? Do they look reasonable in this
    case? Carry out this test and discuss your results.

```{r}
t.test(pim$bp[pim$type=="Yes"],pim$bp[pim$type=="No"], var.equal = TRUE)
t.test(bp ~ type, data = Pima.te)
```

H0: mean1 == mean2 H1: mean1 != mean2 We can use t test. Because std is
unknown. We can reject H0, so the treatment actually had an impact on
the memory score of the patients. 因为样本足够大,
中心极限定理证明这是合理的假设. The test two sample test,

(f) What non-parametric tests would be adequate for parts (d) and (e)?
    Carry this test out and compare your results with the tests in (d)
    and (e).

```{r}
wilcox.test(pim$bp[pim$type=="Yes"], mu=70)
wilcox.test(pim$bp[pim$type=="Yes"],pim$bp[pim$type=="No"], alternative = 'less')
#答案
wilcox.test(Pima.te$bp, mu = 70)
wilcox.test(bp ~ type, data = Pima.te) # 不用写一堆条件. 
```

The result is same with (d) But the result is not same with (e),so the
treatment doesn't had an impact on the memory score of the patients.
答案: The p-value in this case is 0.0173, which is higher than before.
At the 5% or 2% level we would reach the same conclusion as in (d), but
at the 1% level it would be different The p-value is small and the
conclusion coincides with the t-test.
