---
title:    |
  | STAT 210
  | Applied Statistics and Data Analysis: 
  | Homework 6
author: Juyi Lin
date: "Due on Oct. 31/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
The table below has the results of an experiment run to determine the effect of four different oven temperatures on the density of a certain type of ceramics

1. Do a complete analysis of variance for this set. Plot the data. 
2. Write the equation for the model. 
3. Determine whether the treatments have an effect on the amount of hemoglobin in blood by means of a hypothesis test.
4. Plot the diagnostic charts and comment on them. Use also Levene’s test and Shapiro-Wilk. Use Tukey’s HSD procedure to make pairwise comparisons and comment on the results.

```{r}
density <- c(21.8, 21.9, 21.7, 21.6, 21.7,
21.5, 21.4, 21.5, 21.4, 21.6,
21.7, 21.8, 21.8, 21.6, 21.5,
21.9, 21.7, 21.8, 21.6, 21.7)
#temp <- factor(rep(c(100,125,150,175), each = 5))
#Q1data <- data.frame(temp,density)
#str(Q1data)
#model1 <- aov(temp ~ density, data = Q1data)
#plot(yield~soil,col='lightblue3')
```

 
## Question 2
In an experiment to study the effect of fertilizers on the spear elongation in asparagus, four different fertilizers
and a control group (no fertilizer) were tested and five asparagus spears were measured for each treatment.
The treatments are coded trmt1, trmt2, trmt3, trmt4, and the control Ctrl. The measurements (length)
is the length in mm of the asparagus spear. The data is in the file spear.
Do a complete analysis of variance for this set. Plot the data. Write the equation for the model. Determine
whether the treatments have an effect of the length of the asparagus spear by means of a hypothesis test.
Plot the diagnostic charts and comment on them. Use also Levene’s test and Shapiro-Wilk. Use Tukey’s
HSD procedure to make pairwise comparisons and comment on the results.

```{r}

```


