---
title: "Hw6-2022"
output: html_document
date: "2022-10-25"
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question1

The table below has the results of an experiment run to determine the
effect of four different oven temperatures on the density of a certain
type of ceramics

1.  Do a complete analysis of variance for this set. Plot the data.

2.  Write the equation for the model.

3.  Determine whether the treatments have an effect on the amount of
    hemoglobin in blood by means of a hypothesis test.

4.  Plot the diagnostic charts and comment on them. Use also Levene's
    test and Shapiro-Wilk. Use Tukey's HSD procedure to make pairwise
    comparisons and comment on the results.

```{r}
density <- c(21.8, 21.9, 21.7, 21.6, 21.7,
21.5, 21.4, 21.5, 21.4, 21.6,
21.7, 21.8, 21.8, 21.6, 21.5,
21.9, 21.7, 21.8, 21.6, 21.7)
temp <- factor(rep(c(100,125,150,175), each = 5))
Q1data <- data.frame(temp,density)
str(Q1data)
```

From the formula for SST , we can obtain the total sum of squares by
finding the differences between the data and the overall mean:

```{r}
sum((density-mean(density))^2)
```

```{r}
boxplot(density~temp, data = Q1data) #直接plot也可以
points(density~temp, data = Q1data, pch=16,col=4)
```
答案:
人们注意到，这些盒子的大小相似，这表明方差是相等的。温度为125◦的密度似乎比其他温度值要低。


```{r}
mod1 = aov(density~temp, data = Q1data)
summary(mod1)
anova(mod1)
print(mod1$coefficients)
```

p = 0.0058, alpha = 0.01, p \< alpha, reject null hypothesis. We choose
alternative that at least two of means are different. The treatments
have an effect.

anova可以看出,The p-value is small and we reject the null hypothesis of no effect.



2.  write the equation μˆ + τˆ1 = 2.174000e+01 ; μˆ+τˆ2= 2.174000e+01
    -2.600000e-01 = 1.914 μˆ+τˆ3 = 2.174000e+01 μˆ+τˆ4 = 2.174000e+01

Plot the diagnostic charts and comment on them.

```{r}
par(mfrow=c(2,2))
plot(mod1,  cex.lab=0.8, cex.sub=0.8)
```

Since we have 5 replications for each treatment level and there are only
5 x-values, the points appear vertically aligned at these values. We
look in this graphs for constant variance. We see that values in some
cases appear to be more spread than in others, and this may be a sign of
non-constant variance. However, we only have a few points and this is
difficult to determine.
答案是说没有问题, 看起来很好, qqplot说明正态性满足. 

Next, we do the Shapiro-Wilk and Levene tests for the residuals.

```{r}
shapiro.test(resid(mod1))
```

This shows that at the 5% level (or lower levels) we cannot reject the
null hypothesis of Gaussianity.答案是normality,

```{r}
library(car)
leveneTest(mod1)
```

This test has a large p-value, saying that hypothesis of homoscedasticy
is not rejected.

Use Turkey's HSD procedure to make pairwise comparisons and comment on
the results.

```{r}
(mod1.tky = TukeyHSD(mod1,conf.level = 0.95))
plot(mod1.tky)
```

We get a table with six comparisons. we can see the value of difference.
The 150-100, 175-100, 175-150,150-125 p value higher than 5%, so that
they don't have significant difference. 125-100, 175-125 p value is
smaller than 1%, they have significant difference.  
(还要总结一下, all involve the 125◦ temperature)
然后就可以得出结论, Therefore, a reasonable conclusion is that temperatures of 100◦, 150◦, and 175◦ produce ceramics with similar densities, while temperature 125◦ produces ceramics with lower densities.

If the intervals include zero, then the difference is not significant.
so that 125-100, 175-125 have significant difference.

## Question2

In an experiment to study the effect of fertilizers on the spear
elongation in asparagus, four different fertilizers and a control group
(no fertilizer) were tested and five asparagus spears were measured for
each treatment. The treatments are coded trmt1, trmt2, trmt3, trmt4, and
the control Ctrl. The measurements (length) is the length in mm of the
asparagus spear. The data is in the file spear.

1.  Do a complete analysis of variance for this set. Plot the data.
2.  Write the equation for the model.\
3.  Determine whether the treatments have an effect of the length of the
    asparagus spear by means of a hypothesis test.
4.  Plot the diagnostic charts and comment on them. Use also Levene's
    test and Shapiro-Wilk. Use Tukey's HSD procedure to make pairwise
    comparisons and comment on the results.

```{r}
results <- read.table('spear',header=T)
attach(results)
str(results)
results$treat = as.factor(treat)
```

From the formula for SST , we can obtain the total sum of squares by
finding the differences between the data and the overall mean:

```{r}
sum((length-mean(length))^2)
```



```{r}
model1 = aov(length~treat)
summary(model1)
print(model1$coefficients)
```

null hypothesis: the treatment means are not significantly different. p
is smaller than 0.01, so we can reject the null hypothesis, the
treatment means are significantly different.

Plot the data.

```{r}
boxplot(length~treat, data = results)
points(length~treat, data = results, pch=25,col=4)
```
答案:  Looking at the boxes, one may think that homoscedasticity may be a problem here. However, looking at the points, we see that the dispersion is similar for all treatments


2.  write the equation μˆ + τˆ1 = 94.14 ; μˆ+τˆ2= 94.14+ 0.90 = 95.04
    μˆ+τˆ3 = 94.14-3.30=90.84 μˆ+τˆ4 = 94.14 + 1.42 = 95.56 μˆ+τˆ5 =
    94.14 + 8.36 =112.50

答案: 
The model for the experiment is
yik =μ+τi +εik
with1≤i≤5,1≤k≤5.  μisthe overall mean,τi representstheeffectoftreatmenti,and εik arethe experimental errors, which are assumed to be independent Gaussian random variables with mean 0 and common varianc   
还有画出 anova 别忘了.

Look at diagnostic plots to check the assumptions of the model.
```{r}
plot(model1, which=1, cex.lab=0.8, cex.sub=0.8)
```

Since we have 5 replications for each treatment level and there are only
5 x-values, the points appear vertically aligned at these values. We
look in this graphs for constant variance. We see that values in some
cases appear to be more spread than in others, and this may be a sign of
non-constant variance. However, we only have a few points and this is
difficult to determine.

Use also Levene's test and Shapiro-Wilk.
答案:  the only concern here is whether the residuals are homoscedastic. The quantile plot indicates that the
assumption of normality is satisfied. This is confirmed with a Shapiro-Wilk test:
```{r}
library(car)
shapiro.test(resid(model1))
leveneTest(model1)
```



Shapiro-Wilk shows that at the 5% level (or lower levels) we cannot
reject the null hypothesis of Gaussianity. leveneTest shows that has a
large p-value, saying that hypothesis of homoscedasticy is not rejected.

答案: The p-value is large and we doo not reject the hypothesis of normality. To test for homoscedasticity we use Levene’s test:The p-value is large and homoscedasticity is not rejected. We conclude that the assumptions for the model are satisfied. 

Use Tukey's HSD procedure to make pairwise comparisons and comment on
the results.

```{r}
(model1.tky = TukeyHSD(model1,conf.level = 0.95))
plot(model1.tky)
```

We get a table with ten comparisons. we can see the value of difference.
If we choose alpha = 5%, The trmt4-Ctrl,trmt4-trmt1,trmt4-trmt2 p value
smaller than 5%, so that they have significant difference. Others don't
have significant difference. If we choose alpha = 2%, The trmt4-Ctrl,
trmt4-trmt2 p value smaller than 2%,so that they have significant
difference.Others don't have significant difference. If we choose alpha
= 1%, The trmt4-trmt2 p value smaller than 1%,so that they have
significant difference. Others don't have significant difference.

If the intervals include zero, then the difference is not significant.
From figure we can see that trmt4-Ctrl,trmt4-trmt1,trmt4-trmt2 have
significant difference.

答案: 要做个总结. 
All the significant comparisons involve treatment 4, which is significantly different from the control and from treatments 1 and 2. The difference with treatment 3 has a p-value just above 0.05. From the graph we plotted at the beginning, we see that spear length is larger with treatment 4, and the tests support this conclusion, except for treatment 3.
