---
title: "homework9"
output: html_document
date: "2022-11-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Consider the prostate data set in the faraway package. Consider lpsa as the response variable and exclude the variables svi and gleason from the analysis.

Do an exploratory analysis of the data. Do a matrix of plots. Which variables seem to have a linear relationship with the response? Compute and plot the correlation coefficients for the regressors. Comment on what you obtain.

```{r}
library(faraway)
prostate = faraway::prostate
summary(prostate)
#res <- cor(dataB)
#corrplot.mixed(prostate)

```

(b) Fit a model for lpsa with all the other variables as predictors. Calculate the variance inflation factors and eliminate variables with vif greater than two.

```{r}

```

(c) Starting with the variables selected in (b), do a variable selection procedure using backward elimination with a p to remove equal to 0.15. Do also variable selection using the BIC criterion. Compare the models that you get. Do residual analysis for both of them. Comment on your results.

```{r}

```

(d) Which model would you select and why?

```{r}

```

(e) Suppose a new patient with the following values arrives:

Table 1: Variables for a new patient lcavol lweight age Ibph lcp pgg45 1.44692 3.623 65 0.30 -0.799 15.0 Predict the lpsa for this patient along with appropriate 98% prediction and confidence intervals.

```{r}

```

## Question2

For this question use the data set Birthweight.csv. We will consider only the variables birthwt, mppwt and smoker. They represent the weight of the baby at birth, the weight of the mother before pregnancy and whether the mother smokes, with 1 indicating that the mother is a smoker.'

i)  Subset the data corresponding to the variables mentioned above. Plot birthwt against mppwt and color the dots according to the value of smoker. Add a regression line for birthwt against mppwt. Comment. Print the summary table for the regression and interpret the results.

```{r}
birthweight <- read.csv('Birthweight.csv',header=T)
summary(birthweight)
subset()

```

(ii) We want to add smoker as a categorical regressor to the previous model. Fit a complete model including interaction and work your way to a minimal adequate model. Write down the equation for you final model and interpret the coefficients.

```{r}

```

(iii) Draw a scatter plot of birthwt against mppwt and color the dots according to the value of smoker. Add the regression lines for your model. Predict the birthwt value for a mppwt value of 120 and both values for smoker. Add prediction intervals at the 98% level.

```{r}

```

(iv) State clearly the assumptions on which the regression model is based. Using graphs and hypothesis tests, do a diagnostic analysis for the model you fitted and verify whether these assumptions are satisfied.

```{r}

```
