---
title: "Homework1"
output: html_document
date: "2022-08-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question1

Consider the following system of equations:
```
3x + 2y + 2z + 4w = 28 
2x+ y+ z =14 
2x + 5z + 5w = 28 
6x+2y+2z+ w=37
```
1. Create a matrix in R with the coefficients of the system, and a vector with the constants on the right-hand side of the equations. Call them mat1 and vec1, respectively.
```{r}
mat1 <- matrix(c(3,2,2,4,2,1,1,0,2,0,5,5,6,2,2,1), nrow=4, ncol=4, byrow=TRUE); 
vec1 <- matrix(c(28,14,28,37),nr=4,nc=1)
```
2. Find the inverse of mat1 and call it mat2.
```{r}
mat2 <- solve(mat1)
```
3. Create a list named list1 having as components mat1, vec1, and mat2. Call these components item1,
item2, and item3, respectively.
```{r}
mat2 <- solve(mat1)
```
4. Remove mat1, vec1, and mat2 from the working directory.
```{r}
# rm(mat1,vec1,mat2)
```
5. Solve the system of equations and call the solution vec2.
```{r}
solve(mat1) %*% vec1
```
6. Verify the solution.

7. Verify that if you multiply the inverse matrix mat2 by vec1 you get the solution.
```{r}
mat2 * vec1 
```

8. Find the eigenvectors of mat1 and mat2 and verify that the eigenvectors of mat2 are the reciprocals of the eigenvectors of mat1.
```{r}
eigen(mat1)
eigen(mat2)
```

## Question2

Consider the function f(x) = e−|x|, for x ∈ R. We want to use the MonteCarlo method to estimate the value of the integral

1. Plot a graph of this function in the region where you want to calculate the integral.
```{r pressure, echo=False}
sigmoid <- function(x) exp(-abs(x))
curve(sigmoid,-2,2)
```

2. Generate N = 1000 random numbers with uniform distribution in the rectangle [−2, 2] × [0, 1]. Count how many points fall below the curve f(x) = e−|x| and estimate the integral using the fraction of these points with respect to the total number of points and the area of the rectangle. Call the estimator I1000

```{r}
#x <- runif(1000,-2,2)
#y <- runif(1000,0,1)
#if (x < y){
  
#}
hist(rand.unif, 
     freq = FALSE, 
     xlab = 'x',  
     ylim = c(0, 0.4),
     xlim = c(-3,3),
     density = 20,
     main = "Uniform distribution for the interval [-2,0.8]")
  
```

