---
title: "ProblmList3"
output: html_document
date: "2022-09-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 5

The engineer always sets a level of confidence of 98% for these tests. Find a rejection region for the test at this level. Is the value of the test statistic inside or outside the rejection region? What is your conclusion?

```{r}
 bottles <- c(494.04, 499.07, 497.03, 502.79, 495.57, 498.09, 500.40,
             491.80, 494.34, 498.65)
estmean = mean(bottles)
estvari = var(bottles)
eststanddev = sd(bottles)
```

## 6

 Carry this test out using a command in R and look at the p-value that you obtain. What is your conclusion?
 
```{r}
plot(pressure)
```



## 7
Since the p-value is close to the confidence level that she set, the engineer decides to take a new sample of size 20 and obtains the following values
Using this new sample, repeat the test you carried out in (6) and comment on the results you obtain.

```{r}
bottles2 <- c(497.24, 497.43, 500.64, 490.59, 496.24, 497.44, 501.69,
              489.98, 493.83, 495.60, 504.33, 495.11, 497.94, 495.03,
              490.75, 498.16, 491.87, 492.94, 494.12, 500.08)
```
